---
title: "Bego"
author: "Thomas Huet"
date: "11/12/2020"
header-includes:
  # - \usepackage{float}
  # - \floatplacement{figure}{H}  #make every figure with capti
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 10, fig.height = 10)
# opts_chunk$set(fig.width=12, fig.height=8)
library(kableExtra)
library(dplyr)
library(knitr)
library(magick)
library(leaflet)
library(RPostgreSQL)
library(rdrop2)
library(rgl)
library(plotly)

GHimgs <- "https://github.com/zoometh/C14/tree/main/docs/imgs/"

df.corresp.sect <- data.frame(LabelSect=c("Me","Fo","Vr","Sm","Cs","Vm","Go","Vb"),
                              sectors=c("Les Merveilles","Fontananlba","Valaurette",                                                                      "Sainte-Marie","Valmasque",
                                        "Col du Sabion","Gordolasque","Vei del Bouc"),
                              stringsAsFactors = F)
df.corresp.thm <- data.frame(prem=c("A","C","F","H","N","P","R","T","X"),
                             thm=c("Axes/Tools","Horned fig.","Halberd","Anthropomorph",
                                   "Unfigurative","Dagger","Reticulated/Topographic","Yoke",
                                   "Special fig."),
                             colors=c("green","red","lightgreen","blue","grey",                                                                                  "darkgreen","yellow","orange","darkgrey"),
                             stringsAsFactors = F)
thm.not.selected <- c("A","N","X")
impr.rocks <- list(c("4.3.16D"), # HBZZ
                   c("7.1.8") # CdT
                   )
# rocks - - - - - - - - - - - - - - - - - - - - - 
roches <- read.table("img/ROCHES.csv", sep = ";", header = T, stringsAsFactors = F)
# merge to get full names of sectors
roches <- merge(roches,df.corresp.sect,by="LabelSect",all.x = TRUE)
roches$lbl.r <- paste0(roches$zone,'.',roches$groupe,'.',roches$roche)
# figures  - - - - - - - - - - - - - - - - - - - - - 
figures <- read.table("img/FIGURES.csv", sep = ",", header = T, stringsAsFactors = F)
figures$zone <- as.factor(figures$zone)
figures$prem <- toupper(figures$prem)
figures$lbl.r <- paste0(figures$zone,'.',figures$groupe,'.',figures$roche)
figures <- figures[figures$prem != '',] # rm NA
figures <- figures[!(figures$prem %in% thm.not.selected),] # rm N
figures <- merge(figures,df.corresp.thm,by="prem",all.x = TRUE)
# unique(figures$prem)
# source("func/functions.R") # load/run ? function
# f.lflt.chef(getwd()) # create leaflet html map of the CdT
```

# The site

The site is divided into 7 sectors. Main ones are [**Les Merveilles** sector][] and [**Fontanalba** sector][]

```{r df.Sectors}
sect.roches.count <- roches %>%
  group_by(sectors) %>%
  summarise(nb.rocks=n())
# order
sect.roches.count <- sect.roches.count[with(sect.roches.count, order(-nb.rocks)), ]
kable(sect.roches.count,"html") %>%
  kable_styling(full_width = FALSE, position = "center", font_size=12)

```

The entire site is about 900 ha, with the large majority of engraved rocks in the France territory but also a sector in between France and Italy (**Col du Sabion** sector) and a rock in Italy (**Vei del Bouc** sector)  

```{r lf.Sectors}
htmltools::includeHTML(paste0(getwd(),"/img/Bego.html"))
```

There's about 20,000 figuratives pecked engravings

```{r lf.figures, fig.width=7, fig.height=5}
# TODO: color chart by c'colors' field
figures.prem.sum <- figures %>% count(thm, zone)
# figures.prem.sum <- figures.prem.sum[figures.prem.sum$prem != '',] # rm NA
# figures.prem.sum <- figures.prem.sum[!(figures.prem.sum$prem %in% thm.not.selected),] # rm N
# merge to get thm fullname and colors
figures.prem.sum$lbl <- paste0(figures.prem.sum$thm,'<br>n = ',as.character(figures.prem.sum$n))
fig.count <- figures.prem.sum %>%
  plot_ly(x = ~zone,
          y = ~n,
          type = 'bar', 
          text = ~lbl,
          hoverinfo = 'text',
          hovertext = ~lbl,
          name = ~thm,
          color = ~thm) %>%
  layout(yaxis = list(title = 'Count'), barmode = 'stack')
fig.count
```

The Zone 13 (**Valaurette** sector) shows a distinctive profile with almost only topographic figure

### Spatial distribution

#### Altitudes

Engravings are displayed on rocks between ~1990 m to ~2700

```{r altis.rocks, echo=F, fig.width=7, fig.height=20}
# altis
# roches.altis <- subset(roches, select=c(sectors,lbl.r,Z))
# roches.altis <- roches.altis[!is.na(roches.altis$Z),]
# roches.altis <- roches.altis[roches.altis$Z != 0,]
# min(roches.altis$Z)
# plot_ly(data = roches.altis,
#         x = ~sectors,
#         y = ~Z,
#         color = ~sectors,
#         type = 'scatter',
#         alpha = .5,
#         mode = "markers"
# )
impr.rocks <- roches[roches$lbl.r %in% unlist(impr.rocks),]
impr.rocks.altis <- subset(impr.rocks, select=c(sectors,nom,lbl.r,Z))
impr.rocks.altis$lbl <- paste0("<i>",impr.rocks.altis$nom,"</i><br>",impr.rocks.altis$lbl.r)
# hyperlink
impr.rocks.altis$text <- paste0("<a href='",getwd(),"/#","4-3-16D","'>A</a>")
impr.rocks.altis$text <- "<a href='https://google.com'>A</a>"

roches.altis <- subset(roches, select=c(sectors,lbl.r,Z))
roches.altis <- roches.altis[!is.na(roches.altis$Z),]
roches.altis <- roches.altis[roches.altis$Z != 0,]
# roches.altis <- sample_n(roches.altis,100)
# min(roches.altis$Z)
plot_ly(data = roches.altis,
        x = ~sectors,
        y = ~Z,
        color = ~sectors,
        type = 'scatter',
        alpha = .5,
        mode = "markers",
        text = ~lbl.r,
        hoverinfo = 'text',
        hovertext = ~lbl.r,
        name = ~sectors) %>% 
  add_trace(x=impr.rocks.altis$sectors,
            y=impr.rocks.altis$Z,
            type = 'scatter',
            mode = "markers",
            # color = "#000000",
            inherit = F,
            showlegend = F) %>%
  # impressive
  add_annotations(x=impr.rocks.altis$sectors, y=impr.rocks.altis$Z,
                  text=impr.rocks.altis$lbl,
                  showarrow=T,
                  arrowhead=.5) %>%
  # .. + hyperlink
  add_trace(
    mode = 'text',
    textfont = list(
      color = 'transparent'
    ),
    showlegend = F,
    inherit = T
  )
```

# Impressive rocks

## 4-3-16D

*Roche de l'Homme aux bras en zigzag* (Zone IV, Groupe III, Roche 16D)

In the Arpette sector lies the *Roche de l'Homme aux bras en zigzag* 

```{r lf.HBZZ}
# load the lf map from img/CdT.html
# knitr::include_url("https://yihui.shinyapps.io/miniUI/")
# knitr::include_app(paste0(getwd(),"/img/m.html"), height = '600px')
htmltools::includeHTML(paste0(getwd(),"/img/ZIVGIIIR16D.html"))
```

## *Stèle du Chef de Tribu* (Zone VII, Groupe I, Roche 8)

In the vallées de Merveilles lies the unique stelae of the site, the so-called *Stèle du Chef de Tribu*


```{r lf.CdT}
# load the lf map from img/CdT.html
# knitr::include_url("https://yihui.shinyapps.io/miniUI/")
# knitr::include_app(paste0(getwd(),"/img/m.html"), height = '600px')
htmltools::includeHTML(paste0(getwd(),"/img/ZVIIGIR8.html"))
```


## Sectors

### **Les Merveilles** sector

This is really important..

### **Fontanalba** sector

This is really important..

## Engravings

### Altitudes

```{r altis.engr, echo=F, fig.width=7, fig.height=20}
# altis
figures.altis <- subset(figures, select=c(lbl.r,figure,thm))
roches.altis <- subset(roches, select=c(sectors,lbl.r,Z))
figures.altis <- merge(figures.altis,roches.altis, by="lbl.r", all.x=TRUE)
figures.altis <- figures.altis[figures.altis$sectors == "Les Merveilles", ]
figures.altis <- figures.altis[!is.na(figures.altis$Z),]
figures.altis$thm <- factor(figures.altis$thm, levels = unique(figures.altis$thm))
# figures.altis <- sample_n(figures.altis,250)
plot_ly(data = figures.altis,
        x = ~thm,
        y = ~Z,
        color = ~thm,
        type = 'scatter',
        alpha = .5,
        mode = "markers"
)
```

## New paragraph

To be continued...
